Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures

### 2012-01-07 ltx/paper.tex
lot of edits. still need to add to inference and results sections. look for next push after dinner

\documentclass[smallcondensed, final]{svjour3}
% \documentclass[smallcondensed, final]{svjour3}[1/1/2012]
\usepackage[square,numbers]{natbib}
\usepackage{amsmath, epsfig}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{easybmat}
\usepackage{footmisc}
\usepackage[usenames,dvipsnames]{color}
\usepackage{subfig}
\renewcommand\algorithmiccomment[1]{// \textit{#1}}
\newcommand{\ignore}[1]{}
\newcommand{\comment}[1]{}
\newcommand{\frank}[1]{\textcolor{red}{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}}}
\newcommand{\willie}[1]{\textcolor{green}{\textsf{\emph{\textbf{\textcolor{green}{#1}}}}}}
\DeclareMathOperator*{\argmax}{arg\,max}

@@

Tracking is related to the topics of video tracking, data association, filtering, point matching, and `tracking of non-rigid objects'. In general, these tasks aim to maintain the localization of the position and/or shape of an object over time. Areas of interest in this field involve developing metrics for judging the similarity between two object representations (for example, potential representations of the same object at adjacent frames in a video), techniques that provide a way to conduct efficient searches over a frame for regions that are similar to a given object, and methods for predicting the future position or appearance of an object. Methods for tracking single, non-rigid, objects are numerous. Color, for example, has long been used for tracking; these techniques often operate by modeling an object with some color-based appearance model, and using this model to find the object in subsequent frames. One classic approach involves modeling the color of a target (in particular, a distribution over the hue-saturation space of a target region) with a Gaussian mixture model (GMM), and choosing subsequent target positions by searching surrounding areas for regions that yield similar GMMs. Furthermore, the GMM is often allowed to adapt over time, and slowly change to model changes in lighting or other time-based variations in the target's color \citep{raja_1998, mckenna_1999, jepson_2003}. Additionally, color has been successfully applied as the feature in a technique known a the mean-shift procedure (or the kernel-based method) for tracking \citep{comaniciu_2003, perez_2002, nummiaro_2003, lee_2011}. This procedure---a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--optimizes the search for the region in a frame which is most similar to a target region in the previous frame. It involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges for each data point. Additionally, the process (in particular, the specification of the neighborhood and the weight-distribution when calculating the weighted mean of nearby data points) is generalized so that multiple kernels can be specified and used to allow for varied clustering behaviors. It can be shown that, through this procedure, each data point becomes associated with a local point of high density (dependent upon the underlying weight distribution specified by the kernel) which allows for clustering \citep{cheng_1995}. The mean shift algorithm has been implemented successfully to allow for a sort of kernel-based object tracking \citep{comaniciu_2003, comaniciu_1999, comaniciu_2000}.

In addition to single object tracking, a great deal of research in this field over the past decade has focused on the development of algorithms to track multiple objects simultaneously. There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera), complex object interactions, object coalescence (where one object is tracked by multiple times/redundantly), objects with similar appearances, variable (and potentially high) numbers of objects, and objects that enter and exit a field of view at different times. Multiple single-object trackers running simultaneously have been shown to be ineffective, as distinct single-object trackers will tend to coalesce and track the same object. To remedy this problem, methods such as \cite{maccormick1999probabilistic} have developed probabilistic principles for maintaining isolation of object trackers. An approach to this problem involving the use of mixture models has also been studied by \cite{vermaak_2003}; this paper found success in maintaining isolation of distinct objects through the use of a nonparametric mixture model.

In the past five years, there has been a push for methods that attempt to provide general algorithms for the detection and tracking of multiple objects in videos. Blob tracking, a basic method attempting to carry out this goal, has been in use for many years, and has found success in arenas where objects are easily discerned from the background of a video and where localization and segmentation of distinct objects is possible. Blob tracking-based methods run into problems when faced with videos where detection is difficult, object appearance or orientation varies heavily, there exists object occlusion, 
\willie{Describe tracking by detection, more on blob tracking, the other few methods i've found for automated joint detection and tracking}

@@

This section provides details on the extracted data, Dirichlet process mixture (DPM) models, dependent Dirichlet process mixture (DDPM) models, and the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM) model. We make explicit the way in which characteristics of video objects can be represented by elements of these models. We formulate the model in a sufficently abstract manner to allow for implementations using specific object representations and extraction procedures in future studies.

@@

Inference performed on the time dependent Dirichlet process mixture model (Section~\ref{sec:inference}) carries out clustering of the data, which isolates each worm-like data cloud and allows each object to be localized and tracked.

@@

The finite mixture model of the previous section can be extended to a Bayesian mixture model by viewing parameters that were previously point values, $\theta_{c_{i}}$ (the mixture component parameters) and $p_{1}, \ldots, p_{K}$ (the mixture component assignment weights), as random variables and providing each with a prior distribution. In this case, the prior distribution $\mathbb{G}_{0}$ is placed on the mixture component parameters, and the prior distribution $\text{Dir}(\alpha/K, \ldots, \alpha/K)$ is placed on the mixture component assignment weights. The resulting Bayesian mixture model can be formulated generatively as
\begin{align}
\begin{split}
\label{bayesian_mixture_model}
	p_{1}, \ldots, p_{K}  &\sim  \text{Dir}(\alpha/K, \ldots, \alpha/K)\\
		\theta_{k}  &\sim  \mathbb{G}_{0}, \hspace{2mm} \forall k \in \{1, \ldots, K\} \\
	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}})
\end{split}
\end{align}
$\forall i \in \{ 1, \ldots, N \}$, where the $x_{i}$ are observations, the $c_{i}$ are the mixture component assignments associated with each observation, the $\theta_{k}$ are parameters defining the $k^{\text{th}}$ mixture component (i.e. the distribution to be mixed, $F(\theta_{k})$), the $\theta_{k}$ are drawn from a prior distribution $\mathbb{G}_{0}$, and $p_{1}, \ldots, p_{K}$ are drawn from a Dirichlet prior parameterized by $\alpha/K, \ldots, \alpha/K$.

@@

A Dirichlet process mixture model, also refered to as an infinite mixture model, is an extension of the Bayesian mixture model described previously. 
When using a DP as a prior in a Bayesian mixture model, $\Theta$ represents the set of parameters of the component mixture distributions. A DPM may be viewed as allowing the prior distribution over the mixture component parameters in a standard mixture model to be distributed according to a DP; this allows for modeling data where the true number of latent mixture components is unknown and arbitrarily large by letting the number of components remain unbounded (note that only a finite number of these components are assigned to the data). In particular, the DPM can be defined generatively as
\begin{align}
\begin{split}
	\mathbb{G} | \alpha, \mathbb{G}_{0}  &\sim  \text{DP}(\alpha, \mathbb{G}_{0}) \\
	\phi_{i} | \mathbb{G}  &\sim  \mathbb{G} \\
	x_{i}|\phi_{i} &\sim F(\phi_{i})
\end{split}
\end{align}
$\forall i \in \{ 1, \ldots, N \}$, where the $x_{i}$ are observations, the $\phi_{i}$ are parameters defining the mixture component from which the $i_{th}$ observation is drawn (i.e. the distribution to be mixed, $F(\phi_{i})$), and the $\phi_{i}$ are drawn from a prior distribution $\mathbb{G}$, which is in turn drawn from a DP with base distribution $\mathbb{G}_{0}$ and parameter $\alpha$. See \cite{gasthaus_2008} and \cite{gasthaus_thesis} for more details on this formulation. Note the difference between the indexing of the clusters in this model and the indexing in the previous two models. This formulation can be shown to be equivalent to the Bayesian mixture model defined in \eqref{bayesian_mixture_model}, when K is taken to be unbounded; as a result, this model is often called an infinite mixture model.

@@

To account for the dependent behavior of data, there has been research on developing models involving a sequence of DPMs, where components of the mixtures are dependent upon (or are sometimes said to be ``tied to'') corresponding components at neighboring positions in the sequence. For example, if the data shows temporal dependence, the goal might be to create a sequence of DPMs, one for each time-step, where the components of the mixture at each step are dependent upon corresponding components in the both the following and previous time steps. \willie{Discuss and cite DDPMs}.

@@

@@

We desire an extraction procedure that is as unsophisticated as possible, both to gauge the robustness of this method on potentially noisy extraction data and to ensure that the procedure is applicable to a wide range of videos. For these reasons we choose the method of frame-differencing for all extraction performed in this paper, which involves recording the positions of pixels (or pixel groups) that have exhibited differences in intensity or value in succesive frames beyond a given threshold. In particular, if we let $I_{t}$ be the image difference obtained by subtracting the value of the $(i,j)^{th}$ pixel in video frame $t$ from the value of the $(i,j)^{th}$ pixel in video frame $t+1$, we can define extraction on frame $t$ to be the process that returns the dataset
\begin{equation}
	\Omega_{t} = \{ (i,j) | I_{t}(i,j) > \phi \}
\end{equation}
where $i$ and $j$ respectively denote the two spatial positions of a pixel, and $\phi$ is a given pixel value threshold. Frame differencing on each frame $t =\{1, \ldots, T \}$ yields the dataset
\begin{equation}
	\Omega = \bigcup_{t=1}^{T} \Omega_{t} = \{ (i,j,t) | I_{t}(i,j) > \phi \}
\end{equation}
which is equivalent to data with the baseline features described in Section~\ref{sec:data}.

Frame differencing is unsophisticated, computationally inexpensive, able to be applied to a wide range of static, single-camera videos (note that videos used in the experiments described in the following sections were chosen to be static; moving-camera videos should be used in conjunction with applicable extraction methods that allow for camera movement). We have found that, when correctly implemented, frame differencing is sufficiently general to extract the desired worm-like data clouds from a variety of videos containing multiple moving objects. A few examples of this extraction procedure on pairs of consecutive video frames can be found in Figure \ref{fig:img_and_framediff}.

@@

\begin{figure}
  \centering
  \subfloat[]{\label{fig:ants_img}\includegraphics[width=0.335\textwidth]{../img/usbh_frame_723.pdf}} 
  \subfloat[]{\label{fig:ants_img2}\includegraphics[width=0.335\textwidth]{../img/usbh_frame_724.pdf}}
  \subfloat[]{\label{fig:ants_img_framediff}\includegraphics[width=0.323\textwidth]{../img/usbh_framediff.png}}\\
  \subfloat[]{\label{fig:traffic2_img}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame1.png}} 
  \subfloat[]{\label{fig:traffic2_img2}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame2.png}}
  \subfloat[]{\label{fig:traffic2_img_framediff}\includegraphics[width=0.3273\textwidth]{../img/traffic2_framediff.png}}\\
  \subfloat[]{\label{fig:pets2009_img}\includegraphics[width=0.333\textwidth]{../img/pets2009_frame1.png}} 
  \subfloat[]{\label{fig:pets2009_img2}\includegraphics[width=0.333\textwidth]{../img/pets2009_frame2.png}}
  \subfloat[]{\label{fig:pets2009_img_framediff}\includegraphics[width=0.333\textwidth]{../img/pets2009_framediff.png}}\\
  %\subfloat[]{\label{fig:longimg}\includegraphics[width=1.05\textwidth]{../img/longimg_6.pdf}} \\
  \subfloat[]{\label{fig:longimg_all}\includegraphics[width=1\textwidth]{../img/longimg_7_detail.pdf}} 
  \caption{Pairs of consecutive frames and the results produced by taking the pixel-wise difference between these frames. \willie{Want to add colorbar that explains how color represents frame number red = frame 1, blue = frame 100}}
  \label{fig:img_and_framediff}
\end{figure}

We furthermore wish to capture the color (or grayscale) value of each pixel and incorporate this color information into our model. For each pixel $x = (i, j, t) \in \Omega$ (defined above), we specify a square, $L$ pixels in length, centered on $(i, j)$, that selects a set of pixels surrounding $\bold{x}$ in frame $t$. We also specify a scalar color value that is able to be computed for each pixel; this value could, for example, be some function of the r-g-b or h-s-v value of a pixel. The color value of each of the selected pixels surrounding $\bold{x}$ is recorded. Afterwards, the set of possible color values (ie the range of color values to which a pixel may be assigned) is partitioned into $V$ bins, and the number of pixels with a color value lying in each of the bins yields a $V$ dimensional vector of `color counts'. We will refer to this extraction technique as `color counting'.

\willie{Show example color-count vector (histogram) for different objects in an image. give more-formal definition like for frame differencing above.}

@@

The frame differencing and color counting extraction outlined in Section~\ref{sec:modelspec_extraction} yields a set of data $\bold{X} \subset \mathbb{R}^{2} \times \mathbb{Z}_{+}^{V} \times \mathbb{Z}_{+}$, where each element $\bold{x} \in \bold{X}$ can be written
\begin{equation}
\bold{x} = ( \bold{x}^{s}, \bold{x}^{c}, t ) = ( x^{s_{1}}, x^{s_{2}}, x^{c_{1}}, \ldots, x^{c_{V}}, t )
\end{equation}
where $\bold{x}^{s} \in \mathbb{R}^{2}$ is the two dimensional vector of spatial coordinates,  $\bold{x}^{c} \in \mathbb{Z}_{+}^{V}$ is the $V$ discrete dimensional vector of color counts, and $t \in \mathbb{Z}_{+}$ is the discrete time index.

In the following experiments, the hue component of a pixel's hue-saturation-value (h-s-v) tuple is recorded from each pixel surrounding a given $\bold{x}^{s}$ in the manner described in Section~\ref{sec:modelspec_extraction} (where we choose $L=5$). Additionally, the set of possible hue values is partitioned into 10 bins, and the number of pixels with a hue value lying in each of the 10 bins is recorded to yield the vector of `color counts' $\bold{x}^{c} = x^{c_{1}}, \ldots, x^{c_{10}}$. Hue is chosen to represent object color since it has been demonstrated in previous work as a simple representation of object appearance that allows for distinct objects to be well differentiated (\willie{cite mckenna, other papers here}).

@@

The distribution families assumed to generate each $\bold{x}$ must be justified. Both the multivariate normal and multinomial distributions were chosen because they are sufficiently simple (and well studied) to allow for tractable inference and sufficiently flexible to provide a reasonable approximation to the data gained during extraction. In particular, the multivariate normal distribution over the spatial features $\bold{x}^{s}$ can be thought to represent the shape of each object as an oval; likewise, data generated by each moving object during extraction are often ovular---as noisy extraction procedures cause some smoothing of edges and corners, producing blob-like shapes even when objects are not particularly round---and centered on a given object. Furthermore, this model is justified as the maximum likelihood parameter estimate of a normal distribution corresponds to the least squares fit of data relative to the mean of the distribution. \willie{needs to be said more accurately and explained why this helps.} Modeling the color features $\bold{x}^{c}$ as draws from a multinomial distribution (equivalently, as draws from a product of discrete distributions), is justified since distinct object tend to generate pixels whose hue values are noisy but yield consistent counts in discrete hue bins.

@@

We wish to formulate a transition kernel $P(\theta_{t} | \theta_{t-1})$ that provides a reasonable representation of how we expect tracked objects to behave. In the interest of formulating a model with the intention of using it to track arbitrary objects, we do not wish to make many assumptions about the long-term behavior of objects (though transition kernels that incorporate, for example, dynamics could potentially be implemented if one is aware beforehand of an object's behavior).

@@

@@

\section{Inference}
\label{sec:inference}
Algorithms that infer latent model parameters in a Bayesian manner given the data gained during extraction provide detection and tracking results. By formulating the detection and tracking task in terms of a generative model, a variety of methods previously developed for statistical inference can be readily applied; we separate presentations of model and inference so that multiple inference methods, each with its own benefits in terms of precision and speed, can be formulated and applied independently to gain detection and tracking results. This section provides background on the two inference procedures implemented in this study, used to perform Bayesian inference on the model described in Section~\ref{sec:modelspec}. The first is a type of Markov Chain Monte Carlo (MCMC) batch inference, which uses Gibbs sampling to generate samples from a posterior distribution over a model, and the second is a type of Sequential Monte Carlo (SMC) inference, also known as a particle filter, which also generates samples from a posterior information, but in this case in a sequential manner that is allowed due to the time-dependent nature of the model.

@@

@@

>>>>

\documentclass[smallcondensed, draft]{svjour3}
% \documentclass[smallcondensed, final]{svjour3}[1/1/2012]
\usepackage[square,numbers]{natbib}
\usepackage{amsmath, epsfig}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{easybmat}
\usepackage{footmisc}
\usepackage[usenames,dvipsnames]{color}
\usepackage{subfig}
\renewcommand\algorithmiccomment[1]{// \textit{#1}}
\newcommand{\ignore}[1]{}
\newcommand{\comment}[1]{}
\newcommand{\frank}[1]{\textcolor{red}{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}}}
\newcommand{\willie}[1]{\textcolor{green}{\textsf{\emph{\textbf{\textcolor{green}{#1}}}}}}
\DeclareMathOperator*{\argmax}{arg\,max}

@@

Tracking is related to the topics of video tracking, data association, filtering, point matching, and `tracking of non-rigid objects'. In general, these tasks aim to maintain the localization of the position and/or shape of an object over time. Areas of interest in this field involve developing metrics for judging the similarity between two object representations (for example, potential representations of the same object at adjacent frames in a video), techniques that provide a way to conduct efficient searches over a frame for regions that are similar to a given object, and methods for predicting the future position or appearance of an object. An overview of object tracking methods can be found in \cite{yilmaz2006object}. Methods for tracking single, non-rigid, objects are numerous. Color, for example, has long been used for tracking; these techniques often operate by modeling an object with some color-based appearance model, and using this model to find the object in subsequent frames. One classic approach involves modeling the color of a target (in particular, a distribution over the hue-saturation space of a target region) with a Gaussian mixture model (GMM), and choosing subsequent target positions by searching surrounding areas for regions that yield similar GMMs. Furthermore, the GMM is often allowed to adapt over time, and slowly change to model changes in lighting or other time-based variations in the target's color \citep{raja_1998, mckenna_1999, jepson_2003}. Additionally, color has been successfully applied as the feature in a technique known a the mean-shift procedure (or the kernel-based method) for tracking \citep{comaniciu_2003, perez_2002, nummiaro_2003, lee_2011}. This procedure---a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--optimizes the search for the region in a frame which is most similar to a target region in the previous frame. It involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges to a number of modes. Additionally, the process (in particular, the specification of the neighborhood and the weight-distribution when calculating the weighted mean of nearby data points) is generalized so that multiple kernels can be specified and used to allow for varied clustering behaviors. It can be shown that, through this procedure, each data point becomes associated with a local point of high density (dependent upon the underlying weight distribution specified by the kernel) which allows for clustering \citep{cheng_1995}. The mean shift algorithm has been implemented successfully to allow for a sort of kernel-based object tracking \citep{comaniciu_2003, comaniciu_1999, comaniciu_2000}.

In addition to single object tracking, a great deal of research in this field over the past decade has focused on the development of algorithms to track multiple objects simultaneously. There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera) \cite{senior2006appearance, cucchiara2004probabilistic, zhou2003background}, complex object interactions \cite{khan_2004, mckenna2000tracking, dockstader2001multiple}, objects with similar appearances \cite{maccormick1999probabilistic, jepson_2003}, variable (and potentially high) numbers of objects \cite{reilly2010detection}, and objects that enter and exit a field of view at different times \cite{stauffer2003estimating, nedrich2010learning}. Multiple single-object trackers running simultaneously have been shown to be ineffective, as distinct single-object trackers will tend to coalesce and track the same object. To remedy this problem, methods such as \cite{maccormick1999probabilistic} have developed probabilistic principles for maintaining isolation of object trackers. An approach to this problem involving the use of a nonparametric mixture model has also found success in maintaining isolation of distinct objects \cite{vermaak_2003}.

Over the past decade, there have been attempts to provide general algorithms for the fully unsupervised detection and tracking of arbitrary objects in videos. Blob tracking, a basic method attempting to carry out this goal, has been used for many years, and has found success in arenas where objects are easily discerned from the background of a video and where localization and segmentation of distinct objects is possible \cite{francois2004real, isard_2001}. Methods involving blob tracking, however, run into problems when faced with videos where detection is difficult, object appearance or orientation varies heavily, and there exists object occlusion \cite{song2005model}. To aid the accuracy of these methods, techniques have been developed for performing extraction and segmentation in a joint manner, incorporating statistical methods for maintaining hypotheses of different numbers of detected objects, and introducing some of the tracking methods described previously to track distinct blobs after they have been segmented \cite{collins2003mean, isard_2001}. Another family of methods related to the task of unsupervised detection and tracking of video objects goes under the heading of video segmentation algorithms---these methods extend single-frame image segmentation to maintain coherence of image segments over time, and have had some success when used for the explicit purpose of detecting and tracking foreground objects of videos \cite{brox2003unsupervised, sista2000unsupervised, wang1998unsupervised}. Other attempts to perform unsupervised detection and tracking include methods for clustering short sequences of positions that can be gained by detecting the motion of objects \cite{brostow2006unsupervised, brox2010object}, which aim to return full-length distinct object tracks, and a graph based method that carries out a similar task using spectral clustering \cite{fragkiadaki2011detection}. Another approach based on a Gaussian mixture model has also been used to cluster data gained from moving objects \cite{pece_2002}; this method also develops heuristics for the initialization and elimination of new tracks. The technique we introduce in this paper falls into this category of methods, as it aims to provide a general unsupervised detection and tracking method for arbitrary objects in videos via clustering. Differing from previous work, we use a type of time dependent Bayesian nonparametric mixture model, and show how it can be applied to a variety of extraction data to perform detection and tracking.

@@

This section provides details on the data gained during extraction procedures, and on Dirichlet process mixture (DPM) models, dependent Dirichlet process mixture (DDPM) models, and the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM) model. We make explicit the way in which characteristics of video objects can be represented by elements of the GPUDDPM. We formulate the model in a sufficently abstract manner to allow for implementations in future studies that may use more-specific object representations or extraction procedures.

@@

Modeling this data with a time dependent Dirichlet process mixture model (Section~\ref{sec:gpuddpm}) allows us to perform inference (Section~\ref{sec:inference}), which carries out clustering of the data. By choosing this type of model, we aim to provide a clustering result that isolates each worm-like data cloud and consequently allows each object to be localized and tracked.

@@

The finite mixture model of the previous section can be extended to a Bayesian mixture model by viewing parameters that were previously point values, $\theta_{c_{i}}$ (the mixture component parameters) and $p_{1}, \ldots, p_{K}$ (the mixture component assignment weights), as random variables and providing each with a prior distribution. In this case, the prior distribution $\mathbb{G}_{0}$ is placed on the mixture component parameters, and the prior distribution $\text{Dir}(\alpha/K, \ldots, \alpha/K)$ is placed on the mixture component assignment weights. The resulting Bayesian mixture model can be formulated generatively as
\begin{align}
\begin{split}
\label{bayesian_mixture_model}
	p_{1}, \ldots, p_{K}  &\sim  \text{Dir}(\alpha/K, \ldots, \alpha/K)\\
	\theta_{1}, \ldots, \theta_{K}  &\sim  \mathbb{G}_{0} \\
	% \theta_{k}  &\sim  \mathbb{G}_{0}, \hspace{2mm} \forall k \in \{1, \ldots, K\} \\
	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}})
\end{split}
\end{align}
$\forall i \in \{ 1, \ldots, N \}$, where the $x_{i}$ are observations, the $c_{i}$ are the mixture component assignments associated with each observation, the $\theta_{k}$ are parameters defining the $k^{\text{th}}$ mixture component (i.e. the distribution to be mixed, $F(\theta_{k})$), the $\theta_{k}$ are drawn from a prior distribution $\mathbb{G}_{0}$, and $p_{1}, \ldots, p_{K}$ are drawn from a Dirichlet prior parameterized by $\alpha/K, \ldots, \alpha/K$.

@@

A Dirichlet process mixture model, also refered to as an infinite mixture model, is an extension of the Bayesian mixture model described previously. 
When using a DP as a prior in a Bayesian mixture model, $\Theta$ represents the set of parameters of the component mixture distributions. A DPM may be viewed as allowing the prior distribution over the mixture component parameters in a standard mixture model to be distributed according to a DP; this allows for modeling data where the true number of latent mixture components is unknown and arbitrarily large by letting the number of components remain unbounded (note that only a finite number of these components are assigned to the data). In particular, the DPM can be defined generatively as
\begin{align}
\begin{split}
	\mathbb{G} | \alpha, \mathbb{G}_{0}  &\sim  \text{DP}(\alpha, \mathbb{G}_{0}) \\
	\phi_{i} | \mathbb{G}  &\sim  \mathbb{G} \\
	x_{i}|\phi_{i} &\sim F(\phi_{i})
\end{split}
\end{align}
$\forall i \in \{ 1, \ldots, N \}$, where the $x_{i}$ are observations, the $\phi_{i}$ are parameters defining the mixture component from which the $i_{th}$ observation is drawn (i.e. the distribution to be mixed, $F(\phi_{i})$), and the $\phi_{i}$ are drawn from a prior distribution $\mathbb{G}$, which is in turn drawn from a DP with base distribution $\mathbb{G}_{0}$ and parameter $\alpha$. See \cite{gasthaus_2008} and \cite{gasthaus_thesis} for more details on this formulation. Note the difference between the indexing of the clusters in this model and the indexing in the previous two models. This formulation can be shown to be equivalent to the Bayesian mixture model defined in \eqref{bayesian_mixture_model}, when K is taken to be unbounded; as a result, this model is sometimes called an infinite mixture model.

@@

To account for the dependent behavior of data, there has been research on developing models involving a sequence of DPMs, where components of the mixtures are dependent upon (or are sometimes said to be ``tied to'') corresponding components at neighboring positions in the sequence. For example, if the data shows temporal dependence, the goal might be to create a sequence of DPMs, one for each time-step, where the components of the mixture at each step are dependent upon corresponding components in the both the following and previous time steps.

@@

%%%%%
% add more info here on other DDPMs, why we chose the GPUDDPM, how it has an ability to model birth and death of clusters (unlike some sort of sophisticated hidden markov model)
%%%%%

@@

We desire an extraction procedure that is as unsophisticated as possible, both to gauge the robustness of this method on potentially noisy extraction data and to ensure that the procedure is applicable to a wide range of videos. For these reasons we choose the method of frame-differencing for all extraction performed for experiments in this paper (Section~\ref{sec:experiments}), which involves recording the positions of pixels that have exhibited differences in intensity or value in succesive frames beyond a given threshold. In particular, if we let $I_{t}$ be the image difference obtained by subtracting the value of the $(i,j)^{th}$ pixel in video frame $t$ from the value of the $(i,j)^{th}$ pixel in video frame $t+1$, we can define extraction on frame $t$ to be the process that returns the dataset
\begin{equation}
	\Omega_{t} = \{ (i,j,t) \hspace{2mm} | \hspace{2mm} I_{t}(i,j) > \phi \}
\end{equation}
where $i$ and $j$ respectively denote the two spatial positions of a pixel, and $\phi$ is a given pixel value threshold. Frame differencing on each frame $t =\{1, \ldots, T \}$ yields the dataset
\begin{equation}
	\Omega = \bigcup_{t=1}^{T} \Omega_{t} = \{ (i,j,t) \hspace{2mm} | \hspace{2mm} I_{t}(i,j) > \phi,  \hspace{2mm}  \forall t \in \{1, \ldots, T \} \}
\end{equation}
which is equivalent to data with the baseline features described in Section~\ref{sec:data}.

Frame differencing is unsophisticated, computationally inexpensive, able to be applied to a wide range of static, single-camera videos (note that videos used in the experiments described in the following sections were chosen to be static; moving-camera videos should be used in conjunction with applicable extraction methods that provide accurate foreground segmentation for non-stationary video). We have found that, when correctly implemented, frame differencing is sufficiently general to extract the desired worm-like data clouds from a variety of videos containing multiple moving objects. A few examples of this extraction procedure on pairs of consecutive video frames can be found in Figure \ref{fig:img_and_framediff}.

@@

\begin{figure}
  \centering
  \subfloat[]{\label{fig:ants_img}\includegraphics[width=0.335\textwidth]{../img/usbh_frame_723.pdf}} 
  \subfloat[]{\label{fig:ants_img2}\includegraphics[width=0.335\textwidth]{../img/usbh_frame_724.pdf}}
  \subfloat[]{\label{fig:ants_img_framediff}\includegraphics[width=0.323\textwidth]{../img/usbh_framediff.png}}\\
  \subfloat[]{\label{fig:traffic2_img}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame1.png}} 
  \subfloat[]{\label{fig:traffic2_img2}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame2.png}}
  \subfloat[]{\label{fig:traffic2_img_framediff}\includegraphics[width=0.3273\textwidth]{../img/traffic2_framediff.png}}\\
  \subfloat[]{\label{fig:pets2009_img}\includegraphics[width=0.333\textwidth]{../img/pets2009_frame1.png}} 
  \subfloat[]{\label{fig:pets2009_img2}\includegraphics[width=0.333\textwidth]{../img/pets2009_frame2.png}}
  \subfloat[]{\label{fig:pets2009_img_framediff}\includegraphics[width=0.333\textwidth]{../img/pets2009_framediff.png}}\\
  %\subfloat[]{\label{fig:longimg}\includegraphics[width=1.05\textwidth]{../img/longimg_6.pdf}} \\
  \subfloat[]{\label{fig:longimg_all}\includegraphics[width=1\textwidth]{../img/longimg_7_detail.pdf}} 
  \caption{Pairs of consecutive frames and the results produced by taking the pixel-wise difference between these frames (a - f). The final image shows the result of frame differencing over a sequence of images (from the PETS2009 dataset), where color denotes frame number.}
  \label{fig:img_and_framediff}
\end{figure}

We furthermore wish to capture the color (or grayscale) value of each pixel and incorporate this color information into our model. For each pixel $x = (i, j, t) \in \Omega$ (defined above), we specify a square, $L$ pixels in length, centered on $(i, j)$, that selects a set of pixels surrounding $\bold{x}$ in frame $t$. We also specify a scalar color value that is able to be computed for each pixel; this value could, for example, be some function of the red-green-blue (r-g-b) or hue-saturation-value (h-s-v) characteristic of a pixel. The color value of each of the selected pixels surrounding $\bold{x}$ is recorded. Afterwards, the set of possible color values (ie the range of color values to which a pixel may be assigned) is partitioned into $V$ bins, and the number of pixels with a color value lying in each of the bins yields a $V$ dimensional vector of `color counts'. We will refer to this extraction technique as `color counting'.

% perhaps: Show example color-count vector (histogram) for different objects in an image. give more-formal definition like for frame differencing above.

@@

The frame differencing and color counting extraction outlined in Section~\ref{sec:modelspec_extraction} yields a set of data $\bold{X} \subset \mathbb{R}^{2} \times \mathbb{Z}_{+}^{V} \times \{1, \ldots, T \}$, where each element $\bold{x} \in \bold{X}$ can be written
\begin{equation}
\bold{x} = ( \bold{x}^{s}, \bold{x}^{c}, t ) = ( x^{s_{1}}, x^{s_{2}}, x^{c_{1}}, \ldots, x^{c_{V}}, t )
\end{equation}
where $\bold{x}^{s} \in \mathbb{R}^{2}$ is the two dimensional vector of spatial coordinates,  $\bold{x}^{c} \in \mathbb{Z}_{+}^{V}$ is the $V$ discrete dimensional vector of color counts, and $t \in \{1, \ldots, T \}$ is the discrete time index.

In the following experiments, the hue component of the h-s-v value is recorded from all pixels that surround each extracted pixel $\bold{x}$ in the manner described in Section~\ref{sec:modelspec_extraction} (where we choose $L=5$). Additionally, the set of possible hue values is partitioned into 10 bins, and the number of pixels with a hue value lying in each of the 10 bins is recorded to yield the vector of `color counts' $\bold{x}^{c} = x^{c_{1}}, \ldots, x^{c_{10}}$. Hue is chosen to represent object color since it has been demonstrated in previous work as a simple representation of object appearance that allows for distinct objects to be well differentiated \cite{perez_2002, raja_1998, mckenna_1999}.

@@

The distribution families assumed to generate each $\bold{x}$ must be justified. Both the multivariate normal and multinomial distributions were chosen because they are sufficiently simple (and well studied) to allow for tractable inference and sufficiently flexible to provide a reasonable approximation to the data gained during extraction. In particular, the multivariate normal distribution over the spatial features $\bold{x}^{s}$ can be thought to represent the shape of each object as an oval; likewise, data generated by each moving object during extraction are often ovular---as noisy extraction procedures cause some smoothing of edges and corners, producing blob-like shapes even when objects are not particularly round---and centered on a given object.
Furthermore, this model is justified as the maximum likelihood parameter estimate of a normal distribution corresponds to the least squares fit of data relative to the normal distribution mean; since extraction data produced by an object clusters around the centroid of the object, the estimated normal distribution mean should provide a reasonable reflection of the object's centroid.
Modeling the color features $\bold{x}^{c}$ as draws from a multinomial distribution (equivalently, as draws from a product of discrete distributions), is justified since we have observed that distinct object tend to generate pixels whose hue values are noisy but yield consistent counts in discrete hue bins.

@@

We wish to formulate a transition kernel $P(\theta_{t} | \theta_{t-1})$ that provides a reasonable representation of how we expect tracked objects to move over time. In the interest of formulating a model with the intention of using it to track arbitrary objects, we do not wish to make many assumptions about the long-term behavior of objects. For example, we choose not to adopt a transition kernel that incorporates object dynamics, though a transition kernel that models this or other sophisticated behavioral tendencies could potentially be implemented in future works if one knows that they are consistent characteristics of an object's movement.

@@

% note:
% don't i have to include more to really define the transition kernel?? I feel there is something that is not fully specified.

@@

\section{Inference}
\label{sec:inference}
Algorithms that infer latent model parameters in a Bayesian manner given the data gained during extraction provide detection and tracking results. By formulating the detection and tracking task in terms of a generative model, a variety of methods previously developed for statistical inference can be readily applied; we separate presentations of model and inference so that multiple inference methods, each with its own benefits in terms of precision and speed, can be formulated and applied independently to gain detection and tracking results. This section provides background on the two inference procedures implemented in this study, both used to perform Bayesian inference on the model described in Section~\ref{sec:modelspec}. The first is a type of Markov Chain Monte Carlo (MCMC) batch inference, which uses Gibbs sampling to generate samples from a posterior distribution over a model, and the second is a type of Sequential Monte Carlo (SMC) inference, also known as a particle filter, which also generates samples from a posterior information, but in a sequential manner that operates by making use of the time-dependent nature of the model.

@@

% MCMC inference : sampling cluster parameters
% --------------------------------------------

@@

% experiments section
% -------------------
